Kung Faux Pandas \\ \large Simplifying privacy protection

### 2018-06-04 paper/main.tex
Auto Update from local Sharelatex

The ability of independent investigators to verify results is perhaps \emph{the} most fundamental feature that separates science from other forms intellectual inquiry.  Unfortunately, the mechanisms of communicating scientific results have changed little since the age of enlightenment and are inadequate enabling reproducibility for some modern scientific endeavors such as machine learning.  Perhaps the most important impediment to reproducibility is the need to share data.  This has long been difficult in health care fields, but will presently become difficult for just about everyone.

% two references don't seem to represent a large body of published literature. Change wording? Include more references?
There is a large body of published literature detailing methods for overcoming this difficulty via ``anonymizing'' and ``synthesising'' data \cite{patki_synthetic_2016, choi_generating_2017}, however at present a mechanism to routinely \emph{use} these techniques has not been forthcoming.

% explain faux here? faux data - similar to faux fur or faux leather - ethically more appealing, less expensive to produce or use
% faux data is a broader term that encompasses any data privacy protection method - synthesis, anonymization, grouping, perturbation, etc.
This paper details an open source software library called \emph{Kung Faux Pandas (KFP)} which allows for the modular use of data security methods into the popular Python Pandas data science library. For those who prefer other tools, KFP also provides Structured Query Language (SQL) interface which enables users to arbitrarily query any data within a database, returning to the user a data set which either does not contain any personal information (synthesized) or has been anonymized by standard methods.

@@

 Over the past 18 years, there has been several significant papers clarifying the definition and importance of reproducibility along with the closely related term replicability. Leek and Peng have defined reproducibility as the "ability to recompute data analytic results given an observed dataset and knowledge of the data analysis pipeline," and replicability as "the chance that an independent experiment targeting the same scientific question will produce a consistent result" \cite{leek_opinion_2015}; others such as Drummond use the terms in a reverse fashion \cite{drummond_replicability_2009}. Despite the semantic disagreement, both groups agree that an independent experiment with confirmatory results is the strongest support of any experiment. One early influential paper on the topic of reproducibility in scientific computing details key factors in reproducibility as having: data, input parameters, documentation, software code, and an environment capable of running the provided software code; items that cannot easily nor clearly be communicated by the traditional research paper \cite{schwab_making_2000}. The previously mentioned authors propose that reproducibility should be treated as a minimum required standard that all published research should meet \cite{peng_reproducible_2006}. Specifically in the context of machine learning, this means including details to a level someone else could create the same environment including hardware configuration, data used for all experiments, source code, documentation on data and how to run/configure software, and tests that verify the software runs correctly. Once a result can be reproduced, new researchers can then build upon the methods, gather new data for testing/validation, or discover alternative methods to replicate a result.

% clarify this section; how does HIPAA allow data to be shared? BAA or De-identification are main methods
In the United States, there's been a long-standing regulation described in the Health Insurance Portability and Accountability Act (HIPAA), which mandates strong security on data related to health care and is backed by significant fines and possible criminal charges \cite{hippaviol}.  Thus, even institutions which share data face a steep institutional risk. This makes it virtually impossible to enable independent verification of machine learning (ML) models which often require enormous amounts of training data.

Perhaps more ominously for researchers, on May 25 , 2018, the European Union's (EU) General Data Protection Regulation (GDPR) has come into full effect \cite{gdpr}, establishing a new paradigm in data ``ownership'' for that jurisdiction.   In short, this set of rules defines the ``owners'' of data to be people from whom it was collected and obliges anyone possessing  that data to  honor the owners' preferences about how the data is kept and used, including a requirement to deletion of any data at the request of its owner at any time in the future.

These rules are ultimately good for citizens, but they create a pickle for scientists studying humans as it's very difficult to follow normal processes in scientific rigor while complying with privacy protection rules.

@@

@@

% some text about goal to make our research reproducible/replicable - provide source code, pip/conda environment files, docker container
% where to put this text? "Unit testing is important but often overlooked - unit tests can show how a researcher validated their code was giving expected results which can often lead to insights not normally communicated in a results section of a journal article."

@@

\subsection{System Structure}

@@

\subsection{Included Methods}

>>>>

The ability of independent investigators to verify results is perhaps \emph{the} most fundamental feature that separates science from other forms intellectual inquiry. Unfortunately, the mechanisms of communicating scientific results have changed little since the age of enlightenment and are inadequate for enabling reproducibility in some modern scientific endeavors such as machine learning. Perhaps the most important impediment to reproducibility is the need to share and have access to data. This has long been difficult in health care fields, but will presently become difficult for just about everyone.

There is a large body of published literature detailing methods for overcoming this difficulty related to data via de-identification (removing key identifiers from data, grouping individuals, and similar anonymization techniques \cite{hippapro}) and synthesizing data (creating data through algorithmic means, population level statistics, randomization, and ideally preserving inter-relationships between records \cite{walonoski_synthea_2018, patki_synthetic_2016, choi_generating_2017}), however at present a mechanism to routinely \emph{use} these techniques has not been forthcoming.

This paper details an open source software library called \emph{Kung Faux Pandas (KFP)} which allows for the modular combination of published privacy protection methods with the popular Python Pandas data science library\cite{mckinney-proc-scipy-2010}. KFP also provides a Structured Query Language (SQL) interface which enables users to arbitrarily query any data within a database, returning to the user a data set which either does not contain any personal information though synthetic means or anonymization. We collectively refer to de-identified data and synthetic data as \emph{faux data}. Faux in it's standard sense of artificial or fake, but more specifically as it is used in the fashion sense of the word: the use of man made materials that look and feel like real leather, fur, etc., but wihout the problematic derivation from live animals and reliance upon limited resources.

@@

% shorten this? Is this the right place in the background for this?
 Over the past 18 years, there has been several significant papers clarifying the definition and importance of reproducibility along with the closely related term replicability. Leek and Peng have defined reproducibility as the "ability to recompute data analytic results given an observed dataset and knowledge of the data analysis pipeline," and replicability as "the chance that an independent experiment targeting the same scientific question will produce a consistent result" \cite{leek_opinion_2015}; others such as Drummond use the terms in a reverse fashion \cite{drummond_replicability_2009}. Despite the semantic disagreement, both groups agree that an independent experiment with confirmatory results is the strongest support of any experiment. One early influential paper on the topic of reproducibility in scientific computing details key factors in reproducibility as having: data, input parameters, documentation, software code, and an environment capable of running the provided software code; items that cannot easily nor clearly be communicated by the traditional research paper \cite{schwab_making_2000}. The previously mentioned authors propose that reproducibility should be treated as a minimum required standard that all published research should meet \cite{peng_reproducible_2006}. Specifically in the context of machine learning, this means including details to a level someone else could create the same environment including hardware configuration, data used for all experiments, source code, documentation on data and how to run/configure software, and tests that verify the software runs correctly. Once a result can be reproduced, new researchers can then build upon the methods, gather new data for testing/validation, or discover alternative methods to replicate a result.

In the health care domain in the United States the key law governing health data is the Health Insurance Portability and Accountability Act (HIPAA). HIPAA is backed by significant civil penalties as well as specifics about security and what can and cannot be done with protected health information \cite{hippaviol}. There are two primary methods by which data can be shared or made public: de-identifcation, the removal of specific identifiers or summarization to a level at which individual identification is minimized, and through strict legal relationships called a business associate agreement. Even institutions which share data within the confines of the laws face a steep institutional risk. These legal barriers makes it difficult to enable independent verification of machine learning (ML) models which often require enormous amounts of training data.

Perhaps more ominously and broadly relevant for researchers, on May 25 , 2018 the European Union's (EU) General Data Protection Regulation (GDPR) has come into full effect \cite{gdpr}. This law establishes a new paradigm in data ``ownership'' for that jurisdiction. This set of rules defines the ``owners'' of data to be people from whom it was collected and obliges anyone possessing that data to honor the owners' preferences about how the data is kept and used, including a requirement to deletion of any data at the request of its owner at any time in the future.

% continue re-writing from here
These rules are ultimately good for citizens, but they create a pickle for scientists studying humans as it's very difficult to follow normal processes in scientific rigor while complying with privacy protection rules.

@@

% some text about goal to make our research reproducible/replicable - provide source code, pip/conda environment files, docker container
% why not full synthesize a full dataset? limited to sql query where relevant data is brought together, easier to model 1 table than lots of separate tables. Modelling lots of separate tables requires extensive meta data configuration and a significant amount of computing power to generate...
% where to put this text? "Unit testing is important but often overlooked - unit tests can show how a researcher validated their code was giving expected results which can often lead to insights not normally communicated in a results section of a journal article."

@@

@@

\subsection{System Architecture}

@@

\subsection{Included Synthesis Methods}
